---
title: "01_Data_Setup"
author: "Calen P. Ryan"
date: "6/30/2021"
output: html_document
---

We have data in dozens of files across decades of time. We need the data in the format we can work with. I'll start bringing things together. A key aspect to this analysis is that it has to be **scaleable**, in that we will later expand the sample size and need to make it as easy as possible to do so without massive changes to the code. 



# Load packages
```{r}
library(tidyverse) # great collection of packages for data carpentry, modelling, and visualization
library(haven) # package for loading Stata's .dta files. 
library(sjlabelled) # good for renaming, changing classes, etc. in the piped dplyr mode

here::here()
# sets a placeholder 'here'. Then we refer to our files based on the *relative* not the *absolute* location.
```



# Load maternal/child data, birthdays, and early growth parameters

# Load and inspect 83-86
```{r}

mbirth2 <-read_dta(here::here("Data/zip_mother_1983_86/mbirth2.dta")) %>% # note: here::here calls the here function
  rename_all(tolower)

dim(mbirth2)
```

```{r}
names(mbirth2)
```
Lots of names that are not easy to interpret. We must read the document
```{r}
file.show(here::here("Data/zip_mother_1983_86/mbirth.txt")) # or you can click on it - easier to 'control + f ' to find words

```

Find the variables that relate to growth in this dataset
```{r}
print("nothing in here yet - place holder")
```


Find the variables in the 91 dataset that pertain to age

All of the community data was collected on the same day - I'm going to check all three files and make sure they all look similar. 
```{r}

```

