---
title: "01_Data_Setup"
author: "Calen P. Ryan"
date: "6/30/2021"
output: html_document
---

We have data in dozens of files across decades of time. We need the data in the format we can work with. I'll start bringing things together. A key aspect to this analysis is that it has to be **scaleable**, in that we will later expand the sample size and need to make it as easy as possible to do so without massive changes to the code. 



# Load packages
```{r}
library(tidyverse) # great collection of packages for data carpentry, modelling, and visualization
library(readr)
library(haven) # package for loading Stata's .dta files. 
library(sjlabelled) # good for renaming, changing classes, etc. in the piped dplyr mode
library(zscorer)
library(lubridate)
library(anthro)
library(childsds)
here::here()
# sets a placeholder 'here'. Then we refer to our files based on the *relative* not the *absolute* location.
```

# Load maternal/child data, birthdays, and early growth parameters

# Load and inspect 83-86
```{r}

mbirth2 <-read_dta(here::here("Data/zip_mother_1983_86/mbirth2.dta")) %>% # note: here::here calls the here function
  rename_all(tolower)

mlong <-read_dta(here::here("Data/zip_mother_1983_86/mlong.dta")) %>% # note: here::here calls the here function
  rename_all(tolower)

dim(mbirth2)
dim(mlong)
```
```{r}
names(mbirth2)

```
```{r}
names (mlong)
```


Lots of names that are not easy to interpret. We must read the document
```{r}
file.show(here::here("Data/zip_mother_1983_86/mbirth.txt")) # or you can click on it - easier to 'control + f ' to find words

view(mlong)


```

Find the variables that relate to growth in this dataset
```{r}

growth83_86_birthdata<-mbirth2 %>% 
  select(basebrgy,hhldnubi, basewman, sexchild, outcome, daybirth, mnthbrth, yearbirt, dayintw, mothintw, yearintw, bbweight, weight1, weightak, weightgm, heightcm )


 growth83_86_birthdata <- growth83_86_birthdata %>%  
  mutate(weightak_kg = (weightak/1000)) %>% 
  mutate(birthdate = make_date(yearbirt, mnthbrth, daybirth)) %>% 
  mutate(intwdate = make_date(yearintw, mothintw, dayintw)) %>% 
  mutate(age_days_birthweigh = as.numeric(intwdate - birthdate)) %>%
  filter(outcome != "2" | outcome !="3" ) %>% 
  filter(age_days_birthweigh < 30) %>% 
  filter(age_days_birthweigh > 0) %>% 
  filter(weightak > 0) %>% 
  filter(heightcm > 0) 
 
growth83_86_birthdata <- addWGSR(data = growth83_86_birthdata, sex = "sexchild", firstPart= "weightak_kg", secondPart = "age_days_birthweigh", index = "wfa")

growth83_86_birthdata <- addWGSR(data = growth83_86_birthdata, sex = "sexchild", firstPart = "heightcm", secondPart = "age_days_birthweigh", index = "hfa")

```

```{r}
 mlong_anthro <-mlong %>% 
  select(basebrgy, basewman, weight, height, yearms, mntham, daymes) %>% 
  mutate(date_inf_meas = lubridate::make_date(year = yearms,
                                              month = mntham,
                                              day = daymes)) %>% 
  arrange(desc(basewman, basebrgy, date_inf_meas))

birthdates_data <- growth83_86_birthdata %>% 
  select(basebrgy, basewman, sexchild, birthdate)

growth83_86_infdata <- left_join(birthdates_data, mlong_anthro, by = c("basebrgy", "basewman"))
 
 growth83_86_infdata <- growth83_86_infdata %>% 
   mutate(age_days_infweigh = as.numeric(date_inf_meas - birthdate)) %>% 
   mutate(weight_kg = (weight/1000))


growth83_86_infdata <- addWGSR(data = growth83_86_infdata, sex = "sexchild", firstPart= "weight_kg", secondPart = "age_days_infweigh", index = "wfa")

growth83_86_infdata <- addWGSR(data = growth83_86_infdata, sex = "sexchild", firstPart = "height", secondPart = "age_days_infweigh", index = "hfa")

```

```{r}
growth83_86_birthdata <- growth83_86_birthdata %>% 
  rename( wfaz_birth = wfaz, hfaz_birth = hfaz)

growth83_86_infdata <- growth83_86_infdata %>% 
  rename( wfaz_inf = wfaz, hfaz_inf = hfaz)

growth83_86_data <- left_join(growth83_86_birthdata, growth83_86_infdata, by = c("basebrgy", "basewman", "sexchild", "birthdate"))

growth83_86_data
```

```{r}
ggplot(growth83_86_data, aes(x = weight1))+
  geom_histogram(color = "white")+
  facet_wrap(~sexchild)

ggplot(growth83_86_data, aes(x = weightak))+
  geom_histogram(color = "white")+
  facet_wrap(~sexchild)

ggplot(growth83_86_data, aes(x = weightgm))+
  geom_histogram(color = "white")+
  facet_wrap(~sexchild)
  
ggplot(growth83_86_data, aes(x=heightcm))+
  geom_histogram(color = "white")+
  facet_wrap(~sexchild)

summary_weight83_86 <- growth83_86_data %>%
  group_by(sexchild) %>% 
  drop_na() %>%
  summarize(mean_weight83_86 = mean(weightak),
            std_dev_weight83_86 = sd(weightak),
            med_weight83_86 = median(weightak),
            count = n(), .groups = "keep")

summary_height83_86 <- growth83_86_data %>%
  group_by(sexchild) %>% 
  drop_na() %>%
  summarize(mean_height83_86 = mean(heightcm),
            std_dev_height83_86 = sd(heightcm),
            med_height83_86 = median(heightcm),
            count = n(), .groups = "keep")

summary_weight83_86
summary_height83_86 
```


#Load and inspect 91 data
```{r}
child2 <-read_dta(here::here("Data/zip_child_91/child291.dta")) %>% # note: here::here calls the here function
  rename_all(tolower)

hhold91 <-read_dta(here::here("Data/zip_household_91/hhold91.dta")) %>% # note: here::here calls the here function
  rename_all(tolower)


dim(child2)
dim(hhold91)
```

```{r}
growth2 <-child2 %>%
  select(basebrgy, basewman, basehhno, currbrgy, weghtndx, hightndx) 

intw_date91 <- hhold91 %>% 
  select(basebrgy, basewman, mointr91, dyintrvw, yrintr91)

growth91_data <- left_join(growth2, intw_date91, by = c("basebrgy", "basewman"))
         
growth91_data <- left_join(growth91_data, birthdates_data, by = c("basebrgy", "basewman"))

growth91_data
```

```{r}

growth91_data<- growth91_data %>% 
  mutate(intwdate91 = make_date( yrintr91, mointr91, dyintrvw)) %>% 
  mutate(age91_days = as.numeric(intwdate91 - birthdate)) %>% 
  mutate(age91_years = as.numeric(intwdate91 - birthdate)/365.25) %>% 
  mutate(age_mo_91 = age91_years*12) %>% 
  filter(age91_days > 0) %>% 
  filter(weghtndx > 0) %>% 
  filter(hightndx > 0)

growth91_data <- addWGSR(data = growth91_data, sex = "sexchild", firstPart= "weghtndx", secondPart = "age91_days", index = "wfa")

growth91_data <- addWGSR(data = growth91_data, sex = "sexchild", firstPart = "hightndx", secondPart = "age91_days", index = "hfa")

growth91_data <- growth91_data %>% 
  rename( wfaz_91 = wfaz, hfaz_91 = hfaz)


growth91_data

```
```{r}
ggplot(growth91_data, aes(x= weghtndx))+
  geom_histogram(color = "white")+
  facet_wrap(~sexchild)

ggplot(growth91_data, aes(x=hightndx))+
  geom_histogram(color = "white")+
  facet_wrap(~sexchild)

summary_weight91 <- growth91_data %>%
  group_by(sexchild) %>% 
  drop_na() %>%
  summarize(mean_weight91 = mean(weghtndx),
            std_dev_weight91 = sd(weghtndx),
            med_weight91 = median(weghtndx),
            count = n(), .groups = "keep")

summary_height91 <- growth91_data %>%
  group_by(sexchild) %>% 
  drop_na() %>%
  summarize(mean_height91 = mean(hightndx),
            std_dev_height91 = sd(hightndx),
            med_height91 = median(hightndx),
            count = n(), .groups = "keep")

summary_weight91
summary_height91
```

#Load and inspect 94 data


```{r}

cdietiq <-read_dta(here::here("Data/zip_child_94 2/cdietiq.dta")) %>% # note: here::here calls the here function
  rename_all(tolower)

dim(cdietiq)
names(cdietiq)
```

```{r}

```



```{r}
growth94_data <- cdietiq %>% 
  select(basebrgy, basehhn2,basewman,chldwma2,chldcode, mointrv2, dyintrv2, yrintrv2, weghtnd2, hightnd2) %>% 
  filter(chldcode == "1")

growth94_data <- left_join(growth94_data, birthdates_data, by = c("basebrgy", "basewman")) 

growth94_data <- growth94_data %>% 
  mutate(intwdate3 = make_date(yrintrv2, mointrv2, dyintrv2)) %>% 
  mutate(age3_days = as.numeric(intwdate3 - birthdate)) %>% 
  mutate(age3_years = as.numeric(intwdate3 - birthdate)/365.25) %>% 
  mutate(age3_months = age3_years*12) %>% 
  filter(age3_days > 0) %>% 
  filter(weghtnd2 > 0) %>% 
  filter(hightnd2 > 0) %>% 
  drop_na() 
  
height_94_boys <- growth94_data %>%
  filter(sexchild == "1") %>% 
  mutate(heightz_94 = calcz(age= age3_months, val= hightnd2, whotab=who2007_boys))
  
rlang::last_error()

growth94_data
growth94_data$hightnd2
max(who2007_boys$Month)
max(growth94_data$age3_months)

# this works
with(growth94_data, calcz(age=age3_months[1],val=hightnd2[1], whotab=who2007_boys))
##
height_94_boys <- growth94_data %>%
  filter(sexchild == "1")

dim(height_94_boys)
# [1] 88 16
zzz <- rep(0,88)
boy_height <- as.vector(height_94_boys$hightnd2)
boy_age <- as.vector(height_94_boys$age3_months)
boy_height
boy_age

for(j in 1:88) zzz[j] <- calcz(age=boy_age[j], val=boy_height[j], whotab=who2007_boys)
length(zzz)
zzz


## this does weird things. First 88 are fine, but then there are lots of zeros (hundreds).
for(j in 1:88) zzz[j] <- calcz(age=height_94_boys$age3_months[j],val=height_94_boys$hightnd2[j], whotab=who2007_boys)
unlist(zzz)

 zscores_94 <- weightZ_94 = sds(value = growth94_data$weghtnd2,
                          age =   growth94_data$age3_years,
                          sex =   growth94_data$sexchild, male = "1", female = "2",
                          ref =   who.ref,
                          item =  "weight",
                          type =  "SDS"
                           )
 

growth94_data

  
```

```{r} 

ggplot(growth94_data, aes(x=weghtnd2))+
  geom_histogram(color = "white")+
  facet_wrap(~sexchild)

ggplot(growth94_data, aes(x=hightnd2))+
  geom_histogram(color = "white")+
  facet_wrap(~sexchild)

summary_weight94 <- growth94_data %>%
  group_by(sexchild) %>% 
  drop_na() %>%
  summarize(mean_weight94 = mean(weghtnd2),
            std_dev_weight94 = sd(weghtnd2),
            med_weight94 = median(weghtnd2),
            max = max(weghtnd2),
            min = min(weghtnd2),
            count = n(), .groups = "keep")

summary_height94 <- growth94_data %>%
  group_by(sexchild) %>% 
  drop_na() %>%
  summarize(mean_height94 = mean(hightnd2),
            std_dev_height94 = sd(hightnd2),
            med_height94 = median(hightnd2),
            max = max(hightnd2),
            min = min(hightnd2),
            count = n(), .groups = "keep")

summary_weight94
summary_height94

```







