---
title: "01_Data_Setup"
author: "Calen P. Ryan"
date: "6/30/2021"
output: html_document
---

We have data in dozens of files across decades of time. We need the data in the format we can work with. I'll start bringing things together. A key aspect to this analysis is that it has to be **scaleable**, in that we will later expand the sample size and need to make it as easy as possible to do so without massive changes to the code. 



# Load packages
```{r}
library(tidyverse) # great collection of packages for data carpentry, modelling, and visualization
library(haven) # package for loading Stata's .dta files. 
library(sjlabelled) # good for renaming, changing classes, etc. in the piped dplyr mode
library(lubridate)
here::here()
# sets a placeholder 'here'. Then we refer to our files based on the *relative* not the *absolute* location.
```



# Load maternal/child data, birthdays, and early growth parameters

# Load and inspect 83-86
```{r}

mbirth2 <-read_dta(here::here("Data/zip_mother_1983_86/mbirth2.dta")) %>% # note: here::here calls the here function
  rename_all(tolower)

dim(mbirth2)
```

```{r}
names(mbirth2)
```
Lots of names that are not easy to interpret. We must read the document
```{r}
file.show(here::here("Data/zip_mother_1983_86/mbirth.txt")) # or you can click on it - easier to 'control + f ' to find words

```

Find the variables that relate to growth in this dataset
```{r}
print("nothing in here yet - place holder")
```


Find the variables in the 91 dataset that pertain to age

All of the community data was collected on the same day - I'm going to check all three files and make sure they all look similar. 
```{r}

```

```{r}
meaghan_full <-read_dta(here::here("Data/Misc/Full_cohort_for_Meaghan.dta")) %>% # note: here::here calls the here function
  rename_all(tolower)

names(meaghan_full)[1:1000]

sample_dates <-meaghan_full %>% 
  select(basebrgy, basewman, uncchdid, contains("intrvw91"), contains("icage91"))
```

```{r}
source(here::here("Scripts/Functions.R"))

library(tidyverse)

create_date(df = sample_dates, 
            year = yrintrvw91, 
            month = mointrvw91, 
            day = dyintrvw91)

sample_dates$meg_dates <-lubridate::make_date(year = sample_dates$yrintrvw91, 
            month = sample_dates$mointrvw91, 
            day = sample_dates$dyintrvw91)
```

```{r}
iq_dates <-read_dta(here::here("Data/zip_child_91/iq91.dta")) %>% # note: here::here calls the here function
  rename_all(tolower) %>% 
  select(basebrgy, basewman, yeariq, monthiq, dayiq) %>% 
  mutate(iq_date = lubridate::make_date( yeariq+1900, monthiq, dayiq))



```

```{r}
left_join(sample_dates, iq_dates, by = c("basebrgy", "basewman")) %>% 
  mutate(iq_intrvw_date_diff = iq_date - meg_dates) %>% 
  select(basebrgy, basewman, uncchdid, iq_intrvw_date_diff) 
```

```{r}

growth83_86_data<-mbirth2 %>% 
  select(basebrgy,hhldnubi, basewman, sexchild, outcome, daybirth, mnthbrth, yearbirt, dayintw, mothintw, yearintw, bbweight,weight1, weightak, weightgm, heightcm ) %>% 
  mutate(weightak_kg = (weightak/1000)) %>% 
  mutate(birthdate = make_date(yearbirt, mnthbrth, daybirth)) %>% 
  mutate(intwdate = make_date(yearintw, mothintw, dayintw)) %>% 
  mutate(age_at_weigh_days = as.numeric(intwdate - birthdate)) %>%
  filter(outcome != "2" | outcome !="3" ) %>% 
  filter(age_at_weigh_days < 30)
```



```{r}
growth83_86_data %>% 
  mutate(ic_mom_info = paste(basebrgy, basewman, sep = "_")) %>% 
  mutate(age_weigh_month = time_length(days(age_at_weigh_days), unit = "months")) %>% 
  select(ic_mom_info, sexchild, age_weigh_month, weightak_kg, heightcm) %>% 
  top_n(200) %>% 
  rename(id = ic_mom_info, 
         sex = sexchild, 
         agemons = age_weigh_month,
         weight = weightak_kg, 
         height = heightcm) %>% 
  write_csv(here::here("Output/Data/CPR/sample_for_WHOcurves.csv"))
  
  
```

# Testing code for calculating z-scores in 91 from Milan's data
Starting just after calculating growth91_data


```{r}
growth2 <-child2 %>%
  select(basebrgy, basewman, basehhno, currbrgy, weghtndx, hightndx) 

intw_date91 <- hhold91 %>% 
  select(basebrgy, basewman, mointr91, dyintrvw, yrintr91)

birthdates_data <- growth83_86_data %>% 
  select(basebrgy, basewman, sexchild, birthdate)

growth91_data <- left_join(growth2, intw_date91, by = c("basebrgy", "basewman"))
         
growth91_data <- left_join(growth91_data, birthdates_data, by = c("basebrgy", "basewman"))

growth91_data
```


```{r}
growth91_data<- growth91_data %>% 
  mutate(intwdate2 = make_date( yrintr91 + 1900, mointr91, dyintrvw)) %>%
  mutate(birthdate = birthdate + years(1900)) %>% 
  mutate(age2_days = intwdate2 - birthdate) %>% 
  filter(age2_days > 0) %>% 
  filter(weghtndx > 0) %>% 
  filter(hightndx > 0)


zscores91 <- with(growth91_data, anthro_zscores(
  sex = sexchild, age = sjlabelled::as_numeric(age2_days), weight = weghtndx, lenhei = hightndx))

growth91_data
zscores91
# Ok, doesn't calculate as these children are too old. 

```

# CPR attempt at function to calculate z-scores (Based on above)

```{r}

cdietiq <-read_dta(here::here("Data/zip_child_94 2/cdietiq.dta")) %>% # note: here::here calls the here function
  rename_all(tolower)

dim(cdietiq)
names(cdietiq)
```



```{r}
# The calc function seems to work well, given a number for age, a number for weight. 
# There is no 'sex' variable, (although that's wrapped in whotab). 

# I'll first need the data in the right format. Then the function

# Make a months unit for age
growth91_data <-growth91_data %>% 
  mutate(age_mo_91 = time_length(age2_days, unit = "months"))


# Make a function
calcz <- function(age,val,whotab=who2007_girls){
  i <- which(whotab$Month == age)
  z <- (val - whotab[i,"M"])/whotab[i,"StDev"]
  return(z)
}

# Test function
calcz(age = round(100.86242), val =125.0 , whotab =  who2007_girls)
# OK works but I need to round the months.




zzz <- rep(0,nrow(growth91_data))

for(j in 1:nrow(growth91_data)){
  zzz[j] <- calcz(age=round(growth91_data[j,"age_mo_91"]),
                  val=growth91_data[j,"hightndx"],
                  whotab=who2007_girls)
}



```
# Can't get that working. Try the next function


```{r}
growth94_data <- cdietiq %>% 
  select(basebrgy, basehhn2,basewman,chldwma2,chldcode, mointrv2, dyintrv2, yrintrv2, weghtnd2, hightnd2) %>% 
  filter(chldcode == "1")

growth94_data <- left_join(growth94_data, birthdates_data, by = c("basebrgy", "basewman")) 

growth94_data <- growth94_data %>% 
  mutate(intwdate3 = make_date(yrintrv2, mointrv2, dyintrv2)) %>% 
  mutate(age3_days = as.numeric(intwdate3 - birthdate)) %>% 
  mutate(age3_years = as.numeric(intwdate3 - birthdate)/365.25) %>% 
  filter(age3_days > 0) %>% 
  filter(weghtnd2 > 0) %>% 
  filter(hightnd2 > 0) %>% 
  drop_na() 
  



growth_subset <-growth94_data[1:10,] %>% remove_all_labels() %>% zap_formats()



 zscores_94 <- sds(value = growth_subset$weghtnd2,
                          age =   growth_subset$age3_years,
                          sex =   as_character(growth_subset$sexchild), male = "1", female = "2",
                          ref =   who.ref,
                          item =  "weight",
                          type =  "SDS"
                           )
 

growth94_data
# Nope. 

# The example data
x <- data.frame(height=c(50,100,60,54),
                sex=c("m","f","f","m"),
                age=c(0,2.9,0.6,0.2))
sds(value = x$height, age = x$age, sex = x$sex, male = "m", female = "f",
    ref = who.ref, item = "height")
# Ok, works. 
# 
# 

# Try a subset
growth_subset2 <-growth_subset %>% 
  select(weghtnd2, sexchild, age3_years) %>% 
  mutate(sexchild = as_character(sexchild))

# Only works when I reduce the age a lot. Seems like 5 years old is the oldest.
sds(value = growth_subset2$weghtnd2,
    age =   c(2, 3, 3, 2, 3, 2, 2, 3, 3, 3),# Made up numbers
    sex =   growth_subset2$sexchild, male = "1", female = "2",
    ref =   who.ref,
    item =  "weight",
    type =  "SDS")



sds(value = growth_subset2$weghtnd2,
    age =   growth_subset2$age3_years,
    sex =   growth_subset2$sexchild, male = "1", female = "2",
    ref =   japanese.ref,
    item =  "weight",
    type =  "SDS")
# Ok also works for kro.ref (older kids?)

```

Ok, so the function works if I include lower ages. What is the WHO reference data
```{r}
who.ref
# I see. The only older age is for dataset 6, which is [1] "sfsubscap - subscapular skinfold (mm)"

```
# Creating Z-score using WHO reference by merging in matching age, M, sd, and calculating myself. 
```{r}
# Make a months unit for age
growth91_data <-growth91_data %>% 
  mutate(age_mo_91 = time_length(age2_days, unit = "months"))


# Make a function
# calcz <- function(age,val,whotab=who2007_girls){
#   i <- which(whotab$Month == age)
#   z <- (val - whotab[i,"M"])/whotab[i,"StDev"]
#   return(z)
# }

growth91_dataz <-left_join(growth91_data %>%
            filter(sexchild == 2) %>% 
            mutate(age_mo_91 = round(age_mo_91)), 
          who2007_girls, 
          by = c("age_mo_91" = "Month"))


growth91_dataz %>% 
  mutate(heightz = (hightndx - M)/StDev)
```


```{r}
mlong_anthro <-mlong %>% 
  select(basebrgy, basewman, weight, height, yearms, mntham, daymes) %>% 
  mutate(date_inf_meas = lubridate::make_date(year = yearms+1900,
                                              month = mntham,
                                              day = daymes)) %>% 
  arrange(desc(basewman, basebrgy, date_inf_meas))
```


```{r}
mlong_anthro %>% 
  ggplot(., aes(x = date_inf_meas, y = height))+
  geom_point(alpha = 0.2)+
  geom_line(alpha = 0.2)
```

# Look for basebrgy/basewman/uncchdid in one file.
```{r}
full_for_meaghan <-haven::read_dta(here::here("Data/Misc/Full_cohort_for_Meaghan.dta"))

brgy_uncchdid_info <-full_for_meaghan %>% 
  select(uncchdid, basebrgy, basewman)

reprostat_females_w_dnam <-read_csv(here::here("Data/Misc/repro_final_original.csv")) %>% 
  select(uncchdid, reprostat, was.preg, trimester)

unc_base_reprostat <-left_join(brgy_uncchdid_info, reprostat_females_w_dnam, by = "uncchdid")

readr::write_csv(unc_base_reprostat, here::here("Output/Data/CPR/unc_base_reprostat.csv"))
```


